### 2018-03-14 paper/3-abstractions.tex
Add busy build example

Our first main idea is the task-description abstraction:
\begin{minted}[xleftmargin=10pt]{haskell}
type Task c k v = @\std{forall}@ f. c f => (k -> f v) -> k -> Maybe (f v)
\end{minted}
This highly-abstracted type\footnote{Readers familiar with \emph{lenses} or
\emph{profunctor optics} might recognise a familiar pattern. We discuss this
in~\S\ref{sec-related-optics}.} is best introduced by an example.
Consider this \Excel spreadsheet:
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{text}
  A1: 10     B1: A1 + A2
  A2: 20     B2: B1 * 2
\end{minted}
\vspace{1mm}
Here cell \cmd{A1} contains the value \cmd{10}, the cell \cmd{B1} contains
the formula \cmd{A1+A2}, etc. We can represent the formulae of this spreadsheet
with the following task description:
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
sprsh1 :: Task Applicative String Integer
sprsh1 fetch "B1" = Just ((+)   <$> fetch "A1" <*> "A2")
sprsh1 fetch "B2" = Just ((* 2) <$> fetch "B1")
sprsh1 _   _      = Nothing
\end{minted}
\vspace{1mm}
Here we instantiate keys \hs{k} with \hs{String}, and values \hs{v} with \hs{Integer}.
(Real spreadsheet cells would contain a wider range of values, of course.)
The task description \hs{sprsh1} embodies all the \emph{formulae} of the spreadsheet,
but not the input values.  Like every \hs{Task}, \hs{sprsh1} is given a
\hs{fetch} function and a key. It pattern-matches on the key to see if it has a
task description (in the \Excel case, a formula) for it. If not, it returns
\hs{Nothing}, indicating the key is an input. If there is a formula in the cell,
it computes the value of the formula, using \hs{fetch} to find the value of any
keys on which it depends.

@@

Given an acyclic task description, the \hs{busy} build system terminates with a
correct result, but it is not a minimal build system (c.f. the
Definition~\ref{def-minimal}); \hs{busy} keeps no track of keys it has already
brought up-to-date, and will therefore busily recompute the same keys again and
again if they have multiple dependents. We will develop much more efficient
build systems in~\S\ref{sec-implementations}.

>>>>

Our first main idea is the task-description abstraction:
\begin{minted}[xleftmargin=10pt]{haskell}
type Task c k v = @\std{forall}@ f. c f => (k -> f v) -> k -> Maybe (f v)
\end{minted}
This highly-abstracted type\footnote{Readers familiar with \emph{lenses} or
\emph{profunctor optics} might recognise a familiar pattern. We discuss this
in~\S\ref{sec-related-optics}.} is best introduced by an example.
Consider this \Excel spreadsheet:
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{text}
  A1: 10     B1: A1 + A2
  A2: 20     B2: B1 * 2
\end{minted}
\vspace{1mm}
Here cell \cmd{A1} contains the value \cmd{10}, the cell \cmd{B1} contains
the formula \cmd{A1+A2}, etc. We can represent the formulae of this spreadsheet
with the following task description:
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
sprsh1 :: Task Applicative String Integer
sprsh1 fetch "B1" = Just ((+)   <$> fetch "A1" <*> fetch "A2")
sprsh1 fetch "B2" = Just ((* 2) <$> fetch "B1")
sprsh1 _   _      = Nothing
\end{minted}
\vspace{1mm}
Here we instantiate keys \hs{k} with \hs{String}, and values \hs{v} with \hs{Integer}.
(Real spreadsheet cells would contain a wider range of values, of course.)
The task description \hs{sprsh1} embodies all the \emph{formulae} of the spreadsheet,
but not the input values.  Like every \hs{Task}, \hs{sprsh1} is given a
\hs{fetch} function and a key. It pattern-matches on the key to see if it has a
task description (in the \Excel case, a formula) for it. If not, it returns
\hs{Nothing}, indicating the key is an input. If there is a formula in the cell,
it computes the value of the formula, using \hs{fetch} to find the value of any
keys on which it depends.

@@

Given an acyclic task description, the \hs{busy} build system terminates with a
correct result, but it is not a minimal build system (c.f. the
Definition~\ref{def-minimal}). Since \hs{busy} has no memory (\hs{i = ()}), it
cannot keep track of keys it has already built, and will therefore busily
recompute the same keys again and again if they have multiple dependents. We
will develop much more efficient build systems in~\S\ref{sec-implementations}.

Despite its limitations, \hs{busy} can easily handle our \hs{sprsh1} example
from the previous section~\S\ref{sec-task}. Below we initialise the store with
\cmd{A1} set to 10 and all other cells set to 20.

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ store  = initialise () $ \key -> if key == "A1" then 10 else 20
@\ghci@ result = busy sprsh1 "B2" store
@\ghci@ getValue result "B1"
30
@\ghci@ getValue result "B2"
60
\end{minted}
\vspace{1mm}

\noindent
As one can see, \hs{busy} correctly built both \cmd{B1} and \cmd{B2}, and in the
right order (if it had built \cmd{B2} before building \cmd{B1}, the result would
have been $20 * 2 = 40$ instead of $(10 + 20) * 2 = 60$). As an example showing
that \hs{busy} is not minimal, imagine that the formula in cell \cmd{B2} was
\cmd{B1~+~B1} instead of \cmd{B1~*~2}. This would lead to calling
\hs{compute}~\hs{"B1"} twice -- once per each occurrence of \cmd{B1} in the
formula.
